---
title: "Ling 12 Final Project"
author: "Zhimei_Chen"
date: '2022-12-07'
output: 
  html_document:
    toc: true
    toc_depth: 4
    toc_float: true
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### Libraries Needed: 
```{r, warning=FALSE,message=FALSE}
library(dbplyr)
library(tidyverse) 
library(ggplot2)
library(corrplot)
library(ISLR)
library(tidymodels)
library(ggpubr)
tidymodels_prefer()
```

### DATA 
#### Credit Card Approval Prediction\
https://www.kaggle.com/datasets/rikdifos/credit-card-approval-prediction?select=application_record.csv \
application_record.csv contains appliers personal information, which I use as features for predicting.\
credit_record.csv records users' behaviors of credit card. \

### read csv and clean
```{r,warning=FALSE}
options(readr.show_col_types = FALSE)
options(tibble.width = Inf) 
library(janitor)
application <- read_csv("application_record.csv")%>% 
  clean_names()
credit <- read_csv("credit_record.csv")%>% 
  clean_names()
head(application)
```

### DATA: application_data

#### Observations
```{r}
cat("Observation number in application: ", nrow(application))
```

#### Trans NA to "Dont wanna tell"
```{r}
application$occupation_type[is.na(application$occupation_type)] <- "Don't wanna tell"
```

#### turn to characters and factors 
```{r,cols.print=18}
application <- application %>% 
  mutate(id = as.character(id),
         code_gender = factor(code_gender),
         flag_own_car = factor(flag_own_car),
         flag_own_realty = factor(flag_own_realty),
         name_income_type = factor(name_income_type),
         name_education_type = factor(name_education_type),
         name_family_status = factor(name_family_status),
         name_housing_type = factor(name_housing_type),
         flag_mobil = factor(flag_mobil),
         flag_work_phone = factor(flag_work_phone),
         flag_phone = factor(flag_phone),
         flag_email = factor(flag_email),
         occupation_type = factor(occupation_type))
head(application)
```


#### Variables
**id**:	Client number\
**code_gender**: Gender [M,F]\
**flag_own_car**: Is there a car [N,Y]\
**flag_own_realty**: Is there a property [N,Y]\
**cnt_children**: Number of children\
**amt_income_total**: Annual income\
**name_income_type**: Income category [Commercial associate, Pensioner, State servant,  Student, Working]\
**name_education_type**: Education level [Academic degree, Higher education, Incomplete higher, Lower secondary, Secondary / secondary special]\
**name_family_status**: Marital status [Civil marriage, Married, Separated, Single / not married, Widow]\
**name_housing_type**: Way of living [Co-op apartment, House / apartment, Municipal apartment, Office apartment, Rented apartment, With parents]\
**days_birth**: Birthday	Count backwards from current day (0), -1 means yesterday\
**days_employed**: Start date of employment	Count backwards from current day(0). If positive, it means the person currently unemployed.\
**flag_mobile**: Is there a mobile phone [0,1]	\
**flag_work_phone**: Is there a work phone [0,1]\
**flag_phone**: Is there a phone	[0,1]\
**flag_email**: Is there an email	[0,1]\
**occupation_type**: Occupation	[Laborers, Core staff, Sales staff, Managers, Drivers, (Other) , NA's]\
**cnt_fam_members**: Family size	\

#### trans inappropriate days of employees to 0 days
```{r}
application$days_employed[application$days_employed==0] ##no one get 0 days
head(application$days_employed[application$days_employed>0])
#Therefore, I think who don't have work has assigned as "36524" days
#trans the number of inappropriate days of employees to 0 days
application$days_employed[application$days_employed == 365243] <- 0
```

### DATA: credit_data
```{r}
head(credit)
summary(credit)
```
There is no missing data in credit.

#### Variables
id: Client number	\
months_balance: Record month	The month of the extracted data is the starting point, backwards, 0 is the current month, -1 is the previous month, and so on\
status: Status (*more in "transfer status"*)


#### Observations
```{r}
cat("Observation number in credit: ",nrow(credit))
```
#### transfer status 
0: 1-29 days past due --> 1\
1: 30-59 days past due --> 2\
2: 60-89 days overdue --> 3\
3: 90-119 days overdue --> 4\
4: 120-149 days overdue --> 5\
5: Overdue or bad debts, write-offs for more than 150 days --> 6\
C: paid off that month --> 0\
X: No loan for the month --> 0
```{r}
credit_needed <- credit
credit_needed$id <- as.character(credit_needed$id)
credit_needed$status<- as.character(credit_needed$status)

credit_needed$status[credit_needed$status == "5"] <- 6
credit_needed$status[credit_needed$status == "4"] <- 5
credit_needed$status[credit_needed$status == "3"] <- 4
credit_needed$status[credit_needed$status == "2"] <- 3
credit_needed$status[credit_needed$status == "1"] <- 2
credit_needed$status[credit_needed$status == "0"] <- 1
credit_needed$status[credit_needed$status == "C"] <- 0
credit_needed$status[credit_needed$status == "X"] <- 0

head(credit_needed)

```

#### turn to factors
```{r,cols.print=18}
credit_needed <- credit_needed %>% 
  mutate(status = factor(status))
head(credit_needed)
```

first year credit
```{r}
first_year_credit <- credit_needed %>%
             arrange(id,months_balance,status) %>%
             group_by(id) %>%
             slice(1:12) %>%
             mutate(month = seq(1:length(id)))
           
head(first_year_credit)
```

### Combine two datasets

#### Example

An example here who, with ID: 5008805, applied credit card and had been approved.
```{r,warning=FALSE}
first_year_credit[first_year_credit$id == 5008805,]
credit[credit$id == 5008805,]
application[application$id == 5008805,]
```

Therefore, we can combine application data with first year, first month credit data
```{r}
length(unique(first_year_credit$id)) 
length(unique(credit$id))
length(unique(application$id))

has_credit <- merge(application, first_year_credit, 
                  by.x = "id", by.y = "id")
cat(length(unique(has_credit$id)),"people has credit card with Info here.")

first_month <- has_credit %>%
  filter(month == min(month))
```

#### Observe 

those had been approved's information (numbers)
```{r}
numchr <- first_month %>%
  ggplot(aes(x = cnt_children)) +
  geom_boxplot(fill="darkturquoise")+
  ggtitle("Amount of children")
income <- first_month %>%
  ggplot(aes(x = amt_income_total)) +
  geom_boxplot(fill="darkturquoise")+
  ggtitle("Annual income")
age <- first_month %>%
  ggplot(aes(x = abs(days_birth)/365))+
  geom_boxplot(fill="darkturquoise")+
  ggtitle("Age")
employedyear <- first_month %>%
  ggplot(aes(x = abs(days_employed)/365))+
  geom_boxplot(fill="darkturquoise")+
  ggtitle("Years been Employeed")
ctfmem <- first_month %>%
  ggplot(aes(x = cnt_fam_members)) +
  geom_boxplot(fill="darkturquoise")+
  ggtitle("Amount of family members")
ggarrange(numchr, income, age, employedyear, ctfmem, 
          nrow = 3,ncol=2,
          labels = c('a', 'b','c','d','e'))
```

those had been approved's information (factors - YES/NO)
```{r}
gender <- first_month %>%
  ggplot(aes(x = code_gender)) +
  geom_bar(fill="darkturquoise", stat = "count") +
  ggtitle("Gender")
iscar <- first_month %>%
  ggplot(aes(x = flag_own_car)) +
  geom_bar(fill="darkturquoise", stat = "count")+
  ggtitle("Has car?")
isrealty <- first_month %>%
  ggplot(aes(x = flag_own_realty)) +
  geom_bar(fill="darkturquoise", stat = "count")+
  ggtitle("Has realty?")
ismobile <- first_month %>%
  ggplot(aes(x = flag_mobil)) +
  geom_bar(fill="darkturquoise", stat = "count")+
  ggtitle("Has mobile?")
isworkphone <- first_month %>%
  ggplot(aes(x = flag_work_phone)) +
  geom_bar(fill="darkturquoise", stat = "count")+
  ggtitle("Has work phone?")
isphone <- first_month %>%
  ggplot(aes(x = flag_phone)) +
  geom_bar(fill="darkturquoise", stat = "count")+
  ggtitle("Has phone?")
isemail <- first_month %>%
  ggplot(aes(x = flag_email)) +
  geom_bar(fill="darkturquoise", stat = "count")+
  ggtitle("Has email?")

ggarrange(gender, iscar, isrealty, ismobile,isworkphone, isphone, isemail,
          nrow = 2,ncol=4,
          labels = c('a', 'b','c','d','e','f','g'))
```

those had been approved's information (factors - REST)
```{r}
inctype <- first_month %>%
  ggplot(aes(x = name_income_type)) +
  geom_bar(fill="darkturquoise", stat = "count") +
  coord_flip() +
  ggtitle("Name Income Type")
inctype
edutype <- first_month %>%
  ggplot(aes(x = name_education_type)) +
  geom_bar(fill="darkturquoise", stat = "count") +
  coord_flip() +
  ggtitle("Name Education Type")
edutype
houtype <- first_month %>%
  ggplot(aes(x = name_housing_type)) +
  geom_bar(fill="darkturquoise", stat = "count") +
  coord_flip() +
  ggtitle("Name Housing Type")
houtype
occtype <- first_month %>%
  ggplot(aes(x = occupation_type)) +
  geom_bar(fill="darkturquoise")+
  coord_flip() +
  ggtitle("Occupation Type")
occtype
```
